---
title: "KUD3D User Guide"
author: "Vinay Udyawer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{KUD3D User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installation

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("vinayudyawer/KUD3D")
```

The `KUD3D` package consists of helper functions to use in association with the `rayshader` package to visualise 3D kernel utilisation distribution and movement patterns of animals monitored using passive or satellite telemetry.

The majority of the 3D bathymetric plotting will be done using the `rayshader` package. If you dont have that already you can install it like so:

```{r, eval=FALSE}
install.packages("rayshader")

# To install the latest version from Github (recommended):
devtools::install_github("tylermorganwall/rayshader")
```

# Usage

Apart from the `rayshader` and `KUD3D` packages lets load other helpful packages

```{r, eval=TRUE}
sapply(c("ks",
         "rgl",
         "raster",
         "tidyverse",
         "lubridate",
         "sf",
         "rayshader",
         "VTrack",
         "KUD3D"),
       require, character.only = TRUE)

```

There are a few example datasets in this package that can help understand how to use the new functions

```{r eval=TRUE}
## Station information
data(statinfo)

## Tag metadata information
data(taginfo)

## Detection data
data(tagdata)

## Bathymetry data for our example data (Sourced from the GBR30 project)
data(bath)
```


We can use the `VTrack` functions to calculate Center of Activity estimates for passively tracked animals (see [here](http://vinayudyawer.github.io/ATT/docs/ATT_Vignette.html) for details)

```{r, eval=TRUE}
ATTdata<-setupData(Tag.Detections = tagdata,
                   Tag.Metadata = taginfo,
                   Station.Information = statinfo,
                   source = "VEMCO")

COAdata <- COA(ATTdata = ATTdata,
               timestep = 60)
```


Lets setup our data so they are all in the correct coordinate projection (from lat/long to meters) so the KUD volumes are in the correct units

```{r, eval=TRUE}
## Assign coordinate reference systems used here
ll<-CRS("+proj=longlat +datum=WGS84")    ## WSG84 geographic coordinate system (lat/long)
utm<-CRS("+init=epsg:32755")             ## Projected coordinate system for the Great Barrier Reef (EPSG 32755; in meters)
```

Lets setup our tagdata

```{r, eval=TRUE}
## Use one tagged individual as an example (ID 13792)

tag_ll <-
  COAdata %>%
  filter(Tag.ID %in% 13792) %>%
  st_as_sf(coords = c("Longitude.coa", "Latitude.coa"),
           crs = 4326)

## transform detection data from lat/long to meters
tag_utm <- st_transform(tag_ll, crs = 32755)
```

Now lets set up our bathymetry data
```{r, eval=T}
## Convert bath data into a raster for the full GBR region
gbr30<-rasterFromXYZ(bath, crs=ll)

## crop our bathymetry to just within the reef our tagged animal was monitored (i.e. Lodestone Reef)
lodestone_ll <- crop(gbr30,
                     statinfo %>%
                       filter(installation_name %in% "Lodestone") %>%
                       st_as_sf(coords=c("station_longitude", "station_latitude"), crs=4326) %>%
                       as_Spatial() %>% extent() + 0.05)

## project bathymetry raster from lat/long to meters
lodestone_utm <-
  gbr30 %>%
  projectRaster(., crs=utm) %>%
  crop(.,
       statinfo %>%
         filter(installation_name %in% "Lodestone") %>%
         st_as_sf(coords=c("station_longitude", "station_latitude"), crs=4326) %>%
         st_transform(crs = 32755) %>%
         as_Spatial() %>% extent() + 5000)

```


Calculate the 3D KUD using the `ks` package

```{r, eval=TRUE}
tag <-
  tag_utm %>%
  as_Spatial() %>% as.tibble() %>%
  transmute(X = coords.x1,
            Y = coords.x2,
            depth = Sensor.Value.coa) %>%
  as.matrix

H.pi <- Hpi(tag, binned = TRUE) * 3
fhat <- kde(tag, H = H.pi)
```

We can use the vol3d function in the `KUD3D` package to calculate volume of KUD voxels


```{r, eval=TRUE}
vol3d(fhat, cont = 50) ## in m3
vol3d(fhat, cont = 95)

```







